---
title: "Prediction using Practical Machine Learning"
author: "Offikial Developer"
date: "April 1, 2016"
output: html_document
---

# Synopsis

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. 

These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. 

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 

In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. 

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

## Data
The training data for this project are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. 

If you use this document for any purpose please cite them as they have been very generous in allowing their data to be used for this project.

## Submission

The goal of this project is to predict the manner in which they did the exercise (for 20 different test cases), using the "classe" variable in the training set. 

### Executive Summary  
This report describes: 
* model (and how it was built)
* cross validation
* the expected out of sample error

# Data-Processing
```{r}
# install.packages("caret")
# 
# install.packages("rpart")
# install.packages("rpart.plot")

# install.packages("(randomForest")
```

```{r}
library(caret)

library(rpart)
library(rpart.plot)

library(randomForest)

set.seed(1)
```

## Download the data
```{r}
urlTraining <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
urlTesting  <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(url(urlTraining), na.strings=c("NA","#DIV/0!",""))
testing  <- read.csv(url(urlTesting), na.strings=c("NA","#DIV/0!",""))
```

## Process the data

clean the data
```{r}
nzv <- nearZeroVar(training)
training <- training[, -nzv]
testing  <- testing[, -nzv]

training <- training[, colSums(is.na(testing)) == 0]
testing  <- testing[, colSums(is.na(testing)) == 0]

raw_classe <- training$classe

unnecessary <- grepl("^X|timestamp|window", names(training))
training    <- training[, !unnecessary]
training    <- training[, sapply(training, is.numeric)]

unnecessary <- grepl("^X|timestamp|window", names(testing))
testing    <- testing[, !unnecessary]
testing    <- testing[, sapply(testing, is.numeric)]

training$classe <- raw_classe
```  


partition the data
```{r}
partition <- createDataPartition(training$classe, p=0.8, list=FALSE)
trainingData  <- training[partition,]
testData      <- training[-partition,]
```


# Data-Analysis

## Analyse the data
Fit the data (using cross validation)
```{r}
fit <- train(classe ~ ., data=trainingData, method="rf", 
             trControl=trainControl(method="cv", number=3, verboseIter=F))
```  

Model the data
```{r}
model <- fit$finalModel
```  
the expected out of sample error: 0.02%

## Prediction 
```{r}
prediction <- predict(fit, newdata=testing)
```  

# Results
```{r}
decisionTree <- rpart(classe ~ ., data=trainingData, method="class")
prp(decisionTree)
```  


